A formal approach but practical approach in train regulation

### 2025-06-20 paper/main.tex
Nettoyage scories

We suppose that the train network can be described as follows. First, we consider a set of $n$ blocks: $\blocks = \{ \bid{1}, \bid{2}, \ldots, \bid{n}\}$. Blocks are oriented: there is a $\dirForward$ end and a $\dirBackward$ one. Let $\directions = \{\dirForward, \dirBackward\}$ be the set of directions.  Second, we consider a set of $m$ trackworks\footnote{For the sake of simplicity, in this paper, we only consider switches. Our work can easily be extended to crossings and double slip switches.}: $\turnouts = \{\sid{1}, \ldots, \sid{m}\}$. Trackworks are characterized by an internal state: switches can be in direct or deviate position (next denoted '\direct' and '\deviate'), a crossing has only one internal position, a double slip switch can be straight through or crossing. In the examples, to simplify the presentation, we use only switches, but our model is not restricted to them.  Let $\internalState$ denotes the set of internal states. A \emph{network configuration} is given by a function $\sigma: \turnouts \to \internalState$. Let $\networkConf$ denotes the set of all network configurations. For instance, for two turnouts $\theta = \{\sid{1}, \sid{2}\}$, the function $\sid{1} \mapsto \deviate, \sid{2} \mapsto \direct$ states that switch $\sid{1}$ is in deviate position while \sid{2} is in direct one. The latter function is also described via an array  $[\deviate, \direct]$ giving states in trackwork order.

@@

\paragraph{Signals.}

At each end of the blocks, there is a signal whose state is \sigred (red) or \siggreen (green). In other words, the status of signals is described by a function $F: \blocks \times \directions \to \{ \sigred, \siggreen\}$. Given some signal status $F$, a block $\bid{id}$ and a direction $\dirFmt{dir}$, $\signalF{\bid{id}}{dir} = \sigred$ means the signal guarding the exit of block $\bid{id}$ in direction $\dirFmt{dir}$ is red, and we expect trains not leaving that block going in this direction\footnote{Our semantics, presented in Section~\ref{sec:formal-model}, ensure such behaviour.}. The signal status is entirely determined by the network configuration and the locking status due to the regulator\todo{Le régulateur n'est pas encore introduit, je propose de déplacer ça plus bas.}.

>>>>

We suppose that the train network can be described as follows. First, we consider a set of $n$ blocks: $\blocks = \{ \bid{1}, \bid{2}, \ldots, \bid{n}\}$. Blocks are oriented: there is a $\dirForward$ (forward) end and a $\dirBackward$ (backward) one. Let $\directions = \{\dirForward, \dirBackward\}$ be the set of directions.  Second, we consider a set of $m$ trackworks\footnote{For the sake of simplicity, in this paper, we only consider switches. Our work can easily be extended to crossings and double slip switches (e.g. crossings have only one internal state and double slip switches have four states).}: $\turnouts = \{\sid{1}, \ldots, \sid{m}\}$. Trackworks are characterized by an internal state: switches can be in direct or deviate position (next denoted '\direct' and '\deviate'). Let $\internalState$ denotes the set of internal states. A \emph{network configuration} is given by a function $\sigma: \turnouts \to \internalState$. Let $\networkConf$ denotes the set of all network configurations. For instance, for two turnouts $\theta = \{\sid{1}, \sid{2}\}$, the function $\sid{1} \mapsto \deviate, \sid{2} \mapsto \direct$ states that switch $\sid{1}$ is in deviate position while \sid{2} is in direct one. The latter function is also described via an array  $[\deviate, \direct]$ giving states in trackwork order.

@@

\paragraph{Signals.}
\todo{Est-ce qu'on met les signaux dans les agents ?}
At each end of the blocks, there is a signal whose state is \sigred (red) or \siggreen (green). In other words, the status of signals is described by a function $F: \blocks \times \directions \to \{ \sigred, \siggreen\}$. Given some signal status $F$, a block $\bid{id}$ and a direction $\dirFmt{dir}$, $\signalF{\bid{id}}{dir} = \sigred$ means the signal guarding the exit of block $\bid{id}$ in direction $\dirFmt{dir}$ is red, and we expect trains not leaving that block going in this direction\footnote{Our semantics, presented in Section~\ref{sec:formal-model}, ensure such behaviour.}. The signal status is entirely determined by the network configuration and the locking status due to the regulator\todo{Le régulateur n'est pas encore introduit, je propose de déplacer ça plus bas.}.
